Small Case: Checking Hitting Times in the Year of Our Lord

### 2025-02-24 small_case.qmd
Small case: add sim results

```{python Imports} 
import numpy as np
import networkx as nx
import igraph as ig

@@

from matplotlib import pyplot as plt
```

@@

We generate a small 5-node directed, weighted network.
```{python}
G = ig.Graph(
  edges = [
    (0, 1),
    (0, 3),
    (1, 2),
    (1, 3),
    (2, 0),
    (3, 2),
    (3, 4),
    (4, 3),
  ],
  directed = True,
  vertex_attrs = {
    'name': list('ABCDE'),
    'capacity': [10, 12, 20, 10, 3],
  },
  edge_attrs = {
    'weight': [
      30, # (0, 1),
      5, # (0, 3),
      5, # (1, 2),
      10, # (1, 3),
      15, # (2, 0),
      12, # (3, 2),
      6, # (3, 4),
      1, # (4, 3),
    ]
  }
)

@@

fig, ax = plt.subplots()
ig.plot(
  G,
  target = ax,
  vertex_label=G.vs['name'],
  vertex_color='lightcyan',
  vertex_size=50,
  edge_label=G.es['weight'],
  edge_background='white',
  edge_label_dist=0.7,
)
```

We compute Q, the transition matrix.
We extract T, the base transition matrix, by allowing defaults of using the vertex indices as the ordering, since we have the sizes attached to the graph.
```{python}
T = ntc.transition_matrix_from_graph(
  graph=G,
  scaling_per_node=G.vs['capacity'],
  global_scaling=28,
  adjacency_attribute='weight'
)
Q = hmk._Q_mat_naive(T)
```

@@

We can then simulate on this naive static network.
We will offload this work to another script, mostly because the simulations take a significant amount of time.

![Script]()

>>>>

```{python Imports} 
import numpy as np
import networkx as nx
import igraph as ig
import polars as pl

@@

from matplotlib import pyplot as plt, colormaps
import seaborn as sns
```

@@

We generate a small 5-node directed, weighted network.
```{python}
G = ig.Graph(
  edges = [
    (0, 1),
    (0, 3),
    (1, 2),
    (1, 3),
    (2, 0),
    (3, 2),
    (3, 4),
    (4, 3),
  ],
  directed = True,
  vertex_attrs = {
    'name': list('ABCDE'),
    'capacity': [100, 120, 200, 100, 30],
  },
  edge_attrs = {
    'weight': [
      300, # (0, 1),
      50, # (0, 3),
      50, # (1, 2),
      100, # (1, 3),
      150, # (2, 0),
      120, # (3, 2),
      60, # (3, 4),
      10, # (4, 3),
    ]
  }
)

@@

fig, ax = plt.subplots()
ig.plot(
  G,
  target = ax,
  vertex_label=G.vs['name'],
  vertex_color='lightcyan',
  vertex_size=50,
  edge_label=G.es['weight'],
  edge_background='white',
  edge_label_dist=1.7,
)
```

We compute Q, the transition matrix.
We extract T, the base transition matrix, by allowing defaults of using the vertex indices as the ordering, since we have the sizes attached to the graph.
```{python}
TIMESPAN = 28
T = ntc.transition_matrix_from_graph(
  graph=G,
  scaling_per_node=G.vs['capacity'],
  global_scaling=TIMESPAN,
  adjacency_attribute='weight'
)
Q = hmk._Q_mat_naive(T)
```

@@

We can then simulate on this naive static network.
We will offload this work to another script, mostly because the simulations take a significant amount of time (and we only want to run it once)

[script](small_case_sims.py)

Similarly, we offload the construction of the metrics.

[script](small_case_sim_analysis.py)

Here, we construct the empirical CDF
```{python}
metrics = pl.read_parquet("small_case/metrics.parquet")
hitting_time_cols = pl.selectors.starts_with('hitting_time_')
aggregated_metrics = metrics.unpivot(
  on=hitting_time_cols,
  index=['nchains','seed',],
  variable_name='target',
  value_name='hitting_time',
)
max_value_plus_one = aggregated_metrics.select(pl.col('hitting_time').max()).item() + 1
cleaned_metrics = aggregated_metrics.fill_nan(max_value_plus_one).fill_null(max_value_plus_one)

ax = sns.ecdfplot(cleaned_metrics, x='hitting_time', hue='nchains', palette=colormaps.get('jet'))

for k, v in ht_numers.items():
  draw_ecdf(v, ax=ax, label=f"N={k}")

ax.set_xscale('log')
ax.set_xlim(left=0.3, right=3e3)
```

We see relatively good agreement here.
<!-- We detected a bug in this sequence, where the silent graceful failure of numpy.random.Generator.multinomial causes a bug where the transition probabilities that were not normalised correctly earlier cause high probability moves to the "last" node. -->
